---
title: "p8105_hw3_qw2282"
author: "Qinyao Wu"
date: "10/5/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#Instal the data
devtools::install_github("p8105/p8105.datasets")
library(p8105.datasets)
library(ggridges)
library(tidyverse)

#Make the output of figure in the same size. 
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

#Set the theme. 
theme_set(theme_bw() + theme(legend.position = "right"))
```

```{r}
data(brfss_smart2010)

brfss_smart2010 = janitor::clean_names(brfss_smart2010) %>% 
  filter(topic == "Overall Health") %>% #Focus on the overall health topic. 
  mutate(response = forcats::fct_relevel(response, c("Excellent","Very good" , "Good", "Fair", "Poor")))

#state observed in seven locations
brfss_smart2010 %>% 
  filter(year == 2002) %>% 
  group_by(locationabbr, locationdesc) %>%
  summarize(n = n()) %>% 
  group_by(locationabbr) %>% 
  summarize(n = n()) %>% 
  filter(n == 7) %>% 
  knitr::kable()
#States are CT, FL, NC
```

In 2002, CT, FL and NC are the states that observed at seven locations. This indicates that a lot of responses were observed in these three states complared to the others. 

```{r}
num_observations = brfss_smart2010 %>% 
  group_by(year, locationabbr) %>% 
  summarize(n = n())

#Make the sphaghetti plot
ggplot(data = num_observations, aes(x = year, y = n, color = locationabbr)) + geom_line() 

```
The sphaghetti plot is shown with colored locations. 


```{r}
summary_ny_excellent = brfss_smart2010 %>% 
  group_by(locationabbr) %>% 
  filter(locationabbr == "NY" & (year == 2002 | year == 2006 | year == 2010)) %>% 
  select(-c(topic, class, question, sample_size, confidence_limit_low:geo_location)) %>% 
  spread(key = response, value = data_value) %>% 
  group_by(year) 
  knitr::kable(summary_ny_excellent)
  

```





```{r}
　mean_response = brfss_smart2010 %>% 
  group_by(year, locationabbr) %>% 
  select(-c(topic, class, question, sample_size, confidence_limit_low:geo_location)) %>% 
  spread(key = response, value = data_value) %>% 

  summarize(mean_excellent = mean(Excellent),
            mean_very_good = mean(`Very good`),
            mean_good = mean(Good),
            mean_fair = mean(Fair),
            mean_poor = mean(Poor)) %>% 
   gather(key = mean_value, value = mean_proportion, mean_excellent:mean_poor)


 ggplot(mean_response, aes(x = mean_proportion, fill = mean_value)) +
  geom_density(alpha = .5) + 
  facet_grid(~mean_value) + 
  viridis::scale_fill_viridis(discrete = TRUE)



```



##Problem 2

```{r}
data(instacart)


```

This data set contains `r nrow(instacart)` and `r ncol(instacart)`.  Here is a list of the variables in this data set `r colnames(instacart)`. Some key variables are product_name, aisle. 


```{r}
instacart %>% 
  group_by(aisle) %>% #134 aisles
  summarize(n = n()) %>% 
  mutate(order_ranking = min_rank(desc(n))) %>% 
  filter(min_rank(desc(n)) < 2)  #Fresh Vegetables
```

```{r}
instacart %>%
  ggplot( aes(x = aisle_id)) + 
  geom_histogram() + 
  labs(
    title = "Aisle distribution plot",
    x = "Aisle ID",
    y = "Number of orders placed"
  )
```

```{r}
instacart %>% 
  group_by(aisle, product_name) %>% 
  filter(aisle == "baking ingredients" | aisle == "dog food care" | aisle == "packaged vegetables fruits") %>% 
  count() %>%
  group_by(aisle) %>% 
  arrange(n) %>% 
  filter(min_rank(desc(n)) < 2) %>% 
  knitr::kable()


time_order = instacart %>% 
  filter(product_name == "Pink Lady Apples" | product_name == "Coffee Ice Cream") %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour = mean(order_hour_of_day)) %>% 
   spread(key = order_dow, value = mean_hour) %>%
  knitr::kable()
  
  
  
```



##Problem 3

```{r}
data(ny_noaa) 

ny_data = ny_noaa %>% 
  janitor::clean_names() %>% 
  rename(snow_in = snow, tmax_f = tmax, tmin_f = tmin, prcp_mm = prcp) %>% 
  separate(date, into = c("Year", "Month", "Day"), sep = "-") %>% 
  janitor::clean_names() 


ny_data %>%  
   group_by(snow_in) %>%
  summarise(n = n()) %>% 
  mutate(temp_ranking = min_rank(desc(n))) %>% 
  filter(min_rank(desc(n)) < 2)  #0 is the most frequently observed value. 
```  

The most commonly observed value for snow fall is zero. This is because most of the time throughout the year, new york do not get any snow. As a result, a snow fall value of 0 mm will be observed most frequently. 
  

```{R}  
ny_jan_jul = ny_data %>% 
  group_by(id, month) %>% 
  mutate(tmax_f = as.numeric(tmax_f, na.rm = TRUE)) %>% 
  mutate(tmin_f = as.numeric(tmin_f, na.rm = TRUE)) %>%
  filter(month == "01" | month == "07") %>%
  summarize(mean_temp = mean((tmax_f + tmin_f)/2)) %>% 
  na.omit(mean_temp) 
  ggplot(ny_jan_jul, aes( x = id, y = mean_tmax)) +
  geom_point() + 
  facet_grid(~month) + 
  viridis::scale_fill_viridis(discrete = TRUE)
  

```

The observable structure of the pattern is that the temperature in all stations in January is near or under zero degrees, while the temperarture in all stations in july is over 20 degrees. There is no obvious outliers that can be observed. 


```{r}

ny_data %>% 
  na.omit(tmin_f) %>% 
  na.omit(tmax_f) %>% 
  mutate(tmax_f = as.numeric(tmax_f, na.rm = TRUE)) %>% 
  mutate(tmin_f = as.numeric(tmin_f, na.rm = TRUE)) %>%
ggplot(aes(x = tmin_f, y = tmax_f)) + 
  geom_hex()


ny_snow = ny_data %>% 
  filter(snow >0, snow < 100) 
  ggplot(ny_snow, aes(x = snow, fill = year)) +
  geom_density(alpha = 0.2)

```

